---
title: "Work with Models"
author: "Kieran Healy"
date: "01-December-2018"
output: radix::radix_article
---

## Load Libraries


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(gapminder)
library(here)
library(tidyverse)
library(socviz)
```

# Work with Models

```{r model1}
p <- ggplot(data = gapminder,
            mapping = aes(x = log(gdpPercap), y = lifeExp))

p + geom_point(alpha=0.1) +
    geom_smooth(color = "tomato", fill="tomato", method = MASS::rlm) +
    geom_smooth(color = "steelblue", fill="steelblue", method = "lm")

p + geom_point(alpha=0.1) +
    geom_smooth(color = "tomato", method = "lm", size = 1.2, 
                formula = y ~ splines::bs(x, 3), se = FALSE)

p + geom_point(alpha=0.1) +
    geom_quantile(color = "tomato", size = 1.2, method = "rqss",
                  lambda = 1, quantiles = c(0.20, 0.5, 0.85))
```

```{r model2}

p <- ggplot(data = gapminder,
            mapping = aes(x = log(gdpPercap), y = lifeExp))

p + geom_point(alpha=0.1) +
    geom_smooth(color = "tomato", fill="tomato", method = MASS::rlm) +
    geom_smooth(color = "steelblue", fill="steelblue", method = "lm")

p + geom_point(alpha=0.1) +
    geom_smooth(color = "tomato", method = "lm", size = 1.2, 
                formula = y ~ splines::bs(x, 3), se = FALSE)

p + geom_point(alpha=0.1) +
    geom_quantile(color = "tomato", size = 1.2, method = "rqss",
                  lambda = 1, quantiles = c(0.20, 0.5, 0.85))
```


## Show several fits at once, with a legend



```{r severalfits}

model_colors <- RColorBrewer::brewer.pal(3, "Set1")
model_colors


p0 <- ggplot(data = gapminder,
            mapping = aes(x = log(gdpPercap), y = lifeExp))

p1 <- p0 + geom_point(alpha = 0.2) +
    geom_smooth(method = "lm", aes(color = "OLS", fill = "OLS")) +
    geom_smooth(method = "lm", formula = y ~ splines::bs(x, df = 3),
                aes(color = "Cubic Spline", fill = "Cubic Spline")) +
    geom_smooth(method = "loess",
                aes(color = "LOESS", fill = "LOESS"))


p1 + scale_color_manual(name = "Models", values = model_colors) +
    scale_fill_manual(name = "Models", values = model_colors) +
    theme(legend.position = "top")
```


## Look inside model objects

```{r, echo=FALSE}
str(gapminder, strict.width = "wrap", nchar.max = 32, vec.len=1)
```


```{r}
out <- lm(formula = lifeExp ~ gdpPercap + pop + continent,
          data = gapminder)
```

```{r}
summary(out)
```


## Generate predictions to graph

```{r}

min_gdp <- min(gapminder$gdpPercap)
max_gdp <- max(gapminder$gdpPercap)
med_pop <- median(gapminder$pop)

pred_df <- expand.grid(gdpPercap = (seq(from = min_gdp,
                                        to = max_gdp,
                                        length.out = 100)),
                       pop = med_pop,
                       continent = c("Africa", "Americas",
                                     "Asia", "Europe", "Oceania"))

dim(pred_df)
head(pred_df)
```


```{r}

pred_out <- predict(object = out,
                    newdata = pred_df,
                    interval = "predict")
head(pred_out)

```

```{r}

pred_df <- cbind(pred_df, pred_out)
head(pred_df)

```


```{r}

p <- ggplot(data = subset(pred_df, continent %in% c("Europe", "Africa")),
            aes(x = gdpPercap,
                y = fit, ymin = lwr, ymax = upr,
                color = continent,
                fill = continent,
                group = continent))

p + geom_point(data = subset(gapminder,
                             continent %in% c("Europe", "Africa")),
               aes(x = gdpPercap, y = lifeExp,
                   color = continent),
               alpha = 0.5,
               inherit.aes = FALSE) + 
    geom_line() +
    geom_ribbon(alpha = 0.2, color = FALSE) +
    scale_x_log10(labels = scales::dollar)

```


## Tidy model objects with broom

```{r}
library(broom)
```

```{r}
out_comp <- tidy(out)
out_comp %>% round_df()

```


```{r}
p <- ggplot(out_comp, mapping = aes(x = term,
                                    y = estimate))

p + geom_point() + coord_flip() 
```


```{r}
out_conf <- tidy(out, conf.int = TRUE)
out_conf %>% round_df()
```

```{r}
out_conf <- subset(out_conf, term %nin% "(Intercept)")
out_conf$nicelabs <- prefix_strip(out_conf$term, "continent")
```


```{r}

p <- ggplot(out_conf, mapping = aes(x = reorder(nicelabs, estimate),
                                    y = estimate, ymin = conf.low, ymax = conf.high))
p + geom_pointrange() + coord_flip() + labs(x="", y="OLS Estimate")

```


### Get observation-level statistics with augment()

```{r}
out_aug <- augment(out)
head(out_aug) %>% round_df()
```


```{r}
out_aug <- augment(out, data = gapminder)
head(out_aug) %>% round_df()
```

```{r}

p <- ggplot(data = out_aug,
            mapping = aes(x = .fitted, y = .resid))
p + geom_point()

```


### Get model-level statistics with glance()

```{r}
glance(out) %>% round_df()
```


```{r}
library(survival)

out_cph <- coxph(Surv(time, status) ~ age + sex, data = lung)
out_surv <- survfit(out_cph)

```


```{r}
out_tidy <- tidy(out_surv)

p <- ggplot(data = out_tidy, mapping = aes(time, estimate))
p + geom_line() +
    geom_ribbon(mapping = aes(ymin = conf.low, ymax = conf.high), alpha = .2)
```

## Grouped analysis and list columns

```{r}
eu77 <- gapminder %>% filter(continent == "Europe", year == 1977)
```

```{r}
fit <- lm(lifeExp ~ log(gdpPercap), data = eu77)
summary(fit)
```

```{r}

out_le <- gapminder %>%
    group_by(continent, year) %>%
    nest()

out_le

```


```{r}
out_le %>% filter(continent == "Europe" & year == 1977) %>% unnest()
```


```{r, echo = FALSE}
old_digits <- getOption("digits")
options(digits = 3)
```


```{r}

fit_ols <- function(df) {
    lm(lifeExp ~ log(gdpPercap), data = df)
}

out_le <- gapminder %>%
    group_by(continent, year) %>%
    nest() %>% 
    mutate(model = map(data, fit_ols)) 

out_le

```


```{r}

fit_ols <- function(df) {
    lm(lifeExp ~ log(gdpPercap), data = df)
}

out_tidy <- gapminder %>%
    group_by(continent, year) %>%
    nest() %>% 
    mutate(model = map(data, fit_ols),
           tidied = map(model, tidy)) %>%
    unnest(tidied, .drop = TRUE) %>%
    filter(term %nin% "(Intercept)" &
           continent %nin% "Oceania")

out_tidy %>% sample_n(5)

```

```{r, echo = FALSE}
options(digits = old_digits)
```

```{r, fig.height=4, fig.width=10, layout = '1-page'}

p <- ggplot(data = out_tidy,
            mapping = aes(x = year, y = estimate,
                          ymin = estimate - 2*std.error,
                          ymax = estimate + 2*std.error,
                          group = continent, color = continent))

p + geom_pointrange(position = position_dodge(width = 1)) +
    scale_x_continuous(breaks = unique(gapminder$year)) + 
    theme(legend.position = "top") +
    labs(x = "Year", y = "Estimate", color = "Continent")
```


## Plot marginal effects

Note that calculating marginal effects can take some time!

```{r}
library(margins)
```


```{r}

gss_sm$polviews_m <- relevel(gss_sm$polviews, ref = "Moderate")

out_bo <- glm(obama ~ polviews_m + sex*race,
              family = "binomial", data = gss_sm)
summary(out_bo)
```


```{r}
bo_m <- margins(out_bo)
summary(bo_m)
```


```{r}
bo_gg <- as_tibble(summary(bo_m))
prefixes <- c("polviews_m", "sex")
bo_gg$factor <- prefix_strip(bo_gg$factor, prefixes)
bo_gg$factor <- prefix_replace(bo_gg$factor, "race", "Race: ")

bo_gg %>% select(factor, AME, lower, upper) 
```


```{r}
p <- ggplot(data = bo_gg, aes(x = reorder(factor, AME),
                              y = AME, ymin = lower, ymax = upper))

p + geom_hline(yintercept = 0, color = "gray80") +
    geom_pointrange() + coord_flip() +
    labs(x = NULL, y = "Average Marginal Effect") 

```


```{r}
pv_cp <- cplot(out_bo, x = "sex", draw = FALSE)

p <- ggplot(data = pv_cp, aes(x = reorder(xvals, yvals),
                              y = yvals, ymin = lower, ymax = upper))

p + geom_hline(yintercept = 0, color = "gray80") +
    geom_pointrange() + coord_flip() +
    labs(x = NULL, y = "Conditional Effect") 

```


## Plots from complex surveys

```{r}
library(survey)
library(srvyr)
```


```{r}
options(survey.lonely.psu = "adjust")
options(na.action="na.pass")

gss_wt <- subset(gss_lon, year > 1974) %>%
    mutate(stratvar = interaction(year, vstrat)) %>%
    as_survey_design(ids = vpsu,
                     strata = stratvar,
                     weights = wtssall,
                     nest = TRUE)
```


```{r svy_withingroup, tidy = FALSE}

out_grp <- gss_wt %>%
    filter(year %in% seq(1976, 2016, by = 4)) %>%
    group_by(year, race, degree) %>%
    summarize(prop = survey_mean(na.rm = TRUE))

out_grp

```

```{r svy_marginals1}

out_mrg <- gss_wt %>%
    filter(year %in% seq(1976, 2016, by = 4)) %>%
    mutate(racedeg = interaction(race, degree)) %>%
    group_by(year, racedeg) %>%
    summarize(prop = survey_mean(na.rm = TRUE))

out_mrg


```

```{r svy_marginals2}

out_mrg <- gss_wt %>%
    filter(year %in% seq(1976, 2016, by = 4)) %>%
    mutate(racedeg = interaction(race, degree)) %>%
    group_by(year, racedeg) %>%
    summarize(prop = survey_mean(na.rm = TRUE)) %>%
    separate(racedeg, sep = "\\.", into = c("race", "degree"))

out_mrg


```


```{r ch-06-svyyears, fig.cap='Weighted estimates of educational attainment for Whites and Blacks, GSS selected years 1976-2016. Faceting barplots is often a bad idea, and the more facets there are the worse an idea it is. With a small-multiple plot the viewer wants to compare across panels (in this case, over time), but this is difficult to do when the data inside the panels are categorical comparisons shown as bars (in this case, education level by group).', layout = '1-page', fig.width = 6, fig.height = 12}

p <- ggplot(data = subset(out_grp, race %nin% "Other"),
            mapping = aes(x = degree, y = prop,
                          ymin = prop - 2*prop_se,
                          ymax = prop + 2*prop_se,
                          fill = race,
                          color = race,
                          group = race))

dodge <- position_dodge(width=0.9)

p + geom_col(position = dodge, alpha = 0.2) +
    geom_errorbar(position = dodge, width = 0.2) +
    scale_x_discrete(labels = scales::wrap_format(10)) +
    scale_y_continuous(labels = scales::percent) +
    scale_color_brewer(type = "qual", palette = "Dark2") +
    scale_fill_brewer(type = "qual", palette = "Dark2") +
    labs(title = "Educational Attainment by Race",
         subtitle = "GSS 1976-2016",
         fill = "Race",
         color = "Race",
         x = NULL, y = "Percent") +
    facet_wrap(~ year, ncol = 2) +
    theme(legend.position = "top")

```


```{r ch-06-svyyears2, fig.cap='Faceting by education instead.', fig.width = 4, fig.height = 9}

p <- ggplot(data = subset(out_grp, race %nin% "Other"),
            mapping = aes(x = year, y = prop, ymin = prop - 2*prop_se,
                          ymax = prop + 2*prop_se, fill = race, color = race,
                          group = race))

p + geom_ribbon(alpha = 0.3, aes(color = NULL)) +
    geom_line() + 
    facet_wrap(~ degree, ncol = 1) +
    scale_y_continuous(labels = scales::percent) +
    scale_color_brewer(type = "qual", palette = "Dark2") +
    scale_fill_brewer(type = "qual", palette = "Dark2") +
    labs(title = "Educational Attainment\nby Race",
         subtitle = "GSS 1976-2016", fill = "Race",
         color = "Race", x = NULL, y = "Percent") +
    theme(legend.position = "top")
```



### Default plots for models

```{r baseplot}
out <- lm(formula = lifeExp ~ log(gdpPercap) + pop + continent, data = gapminder)
```


```{r}
plot(out, which = c(1,2), ask=FALSE)
```

