---
title: "Supplementary Material"
author: "Kieran Healy"
date: "01-December-2018"
output: radix::radix_article
---

## Load Libraries


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(gapminder)
library(here)
library(tidyverse)
library(ggrepel)
library(socviz)
```

# Supplementary Material

### The basics of accessing and selecting things

```{r}
my_numbers <- c(1, 2, 3, 1, 3, 5, 25)
your_numbers <- c(5, 31, 71, 1, 3, 21, 6)
```

```{r}
my_numbers[4]
my_numbers[7]
```

```{r}
my_numbers[2:4]
```

```{r}
my_tb <- tibble(
    mine = c(1,4,5, 8:11),
    yours = c(3,20,16, 34:31))

class(my_tb)
my_tb
```

```{r}
my_tb[3,1] # Row 3 Col 1
my_tb[1,2] # Row 1, Col 2 
```


```{r}
my_tb[3,"mine"] # Row 3 Col 1
my_tb[1,"yours"] # Row 1, Col 2 
```

```{r}
my_tb[3,"mine"] # Row 3 Col 1
my_tb[1,"yours"] # Row 1, Col 2 
```



```{r}
my_tb[,"mine"] # All rows, Col 1
```


```{r}
my_tb[4,] # Row 4, all cols
```


```{r}
my_tb$mine
```


```{r}
out <- lm(mine ~ yours, data = my_tb)

out$coefficients

out$call

out$qr$rank # nested 
```


```{r}
my_tb$ours <- my_tb$mine + my_tb$yours
my_tb
```

### Tidy data

```{r}
knitr::kable(preg, caption="Some untidy data.", booktabs = TRUE)
```

```{r}
knitr::kable(preg2,
             caption="The same data, still untidy, but in a different way.", booktabs = TRUE)
```

```{r}
preg3 <- preg %>% 
  gather(treatment, n, treatmenta:treatmentb) %>%
  mutate(treatment = gsub("treatment", "", treatment)) %>%
    arrange(name, treatment)
knitr::kable(preg3, 
      caption="Tidied data. Every variable a column, every observation a row.", booktabs = TRUE)
```


```{r, echo=FALSE, message=FALSE}
edu
```

```{r, echo=FALSE, message=FALSE, tidy = FALSE}
edu_t <- gather(data = edu,
                key = school,
                value = freq,
                elem4:coll4)
```


```{r, echo=TRUE, message=FALSE, tidy = FALSE}
edu_t <- gather(data = edu,
                key = school,
                value = freq,
                elem4:coll4)

head(edu_t) 
tail(edu_t) 
```

## Common problems reading in data

### Date formats

```{r, fig.caption="A bad date.", fig.cap = "A bad date."}
p <- ggplot(data = bad_date, aes(x = date, y = N))
p + geom_line()
```


```{r, echo = TRUE, fig.width = 5, fig.height = 4, fig.cap = "Still bad."}

bad_date2 <- rbind(bad_date, bad_date)

p <- ggplot(data = bad_date2, aes(x = date, y = N))
p + geom_line()

```


```{r}
# install.packages("lubridate")
library(lubridate)

bad_date$date <- mdy(bad_date$date)
head(bad_date)

```

```{r, fig.cap="Much better.", out.width="100%", fig.height=4, fig.width=6, fig.margin=TRUE}
p <- ggplot(data = bad_date, aes(x = date, y = N))
p + geom_line()
```


### Year-only dates

```{r, eval = FALSE}

url <- "https://cdn.rawgit.com/kjhealy/viz-organdata/master/organdonation.csv"

bad_year <- read_csv(url)
bad_year %>% select(1:3) %>% sample_n(10)

```

```{r}

bad_year <- read_csv(file = "data/organdonation.csv")
bad_year %>% select(1:3) %>% sample_n(10)

```

```{r, fig.cap="Integer year shown with a decimal point.",  fig.height = 4, fig.width = 5}

p <- ggplot(data = bad_year, aes(x = year, y = donors))
p + geom_point()

```


```{r}
bad_year$year <- int_to_year(bad_year$year)
bad_year %>% select(1:3)
```

## Functions

```{r addnum1, echo = FALSE}
add_xy <- function(x = NULL, y = NULL) {
    x + y
}

```

```{r addnum2}

add_xy(x = 1, y = 7)

```

```{r addnum4}
add_xy(x = 5, y = 2)
```


```{r plot-section-function-1, echo=TRUE, tidy = FALSE}

plot_section <- function(section="Culture", x = "Year",
                         y = "Members", data = asasec,
                         smooth=FALSE){
    require(ggplot2)
    require(splines)
    # Note use of aes_string() rather than aes() 
    p <- ggplot(subset(data, Sname==section),
            mapping = aes_string(x=x, y=y))

    if(smooth == TRUE) {
        p0 <- p + geom_smooth(color = "#999999",
                              size = 1.2, method = "lm",
                              formula = y ~ ns(x, 3)) +
            scale_x_continuous(breaks = c(seq(2005, 2015, 4))) +
            labs(title = section)
    } else {
    p0 <- p + geom_line(color= "#E69F00", size=1.2) +
        scale_x_continuous(breaks = c(seq(2005, 2015, 4))) +
        labs(title = section)
    }

    print(p0)
}

```

```{r, fig.cap='Using a function to plot your results.', out.width="50%", fig.width=4, fig.height=3, fig.show = "hold"}

plot_section("Rationality")
plot_section("Sexualities", smooth = TRUE)

```

```{r plot-section-function-2, echo = TRUE, tidy = FALSE}

plot_section <- function(section="Culture", x = "Year",
                         y = "Members", data = asasec,
                         smooth=FALSE, ...){
    require(ggplot2)
    require(splines)
    # Note use of aes_string() rather than aes() 
    p <- ggplot(subset(data, Sname==section),
            mapping = aes_string(x=x, y=y))

    if(smooth == TRUE) {
        p0 <- p + geom_smooth(color = "#999999",
                              size = 1.2, ...) +
            scale_x_continuous(breaks = c(seq(2005, 2015, 4))) +
            labs(title = section)
    } else {
    p0 <- p + geom_line(color= "#E69F00", size=1.2) +
        scale_x_continuous(breaks = c(seq(2005, 2015, 4))) +
        labs(title = section)
    }

    print(p0)
}


```

```{r, fig.cap='Our custom function can now pass arguments along to fit different smoothers to Section membership data.', out.width="50%", fig.width=4, fig.height=3, fig.show = "hold"}

plot_section("Comm/Urban",
             smooth = TRUE,
             method = "loess")
plot_section("Children",
             smooth = TRUE,
             method = "lm",
             formula = y ~ ns(x, 2))

```


## Preparing the county-level maps

```{r, eval = FALSE, echo = TRUE}

# You will need to use install.packages() to install
# these map and GIS libraries if you do not already
# have them.

library(maptools)
library(mapproj)
library(rgeos)
library(rgdal)

us_counties <- readOGR(dsn="data/geojson/gz_2010_us_050_00_5m.json",
                       layer="OGRGeoJSON")

us_counties_aea <- spTransform(us_counties,
                    CRS("+proj=laea +lat_0=45 +lon_0=-100 \
                         +x_0=0 +y_0=0 +a=6370997 +b=6370997 \
                         +units=m +no_defs"))

us_counties_aea@data$id <- rownames(us_counties_aea@data)
```


```{r, eval = FALSE, echo = TRUE}

alaska <- us_counties_aea[us_counties_aea$STATE == "02",]
alaska <- elide(alaska, rotate=-50)
alaska <- elide(alaska, scale=max(apply(bbox(alaska), 1, diff)) / 2.3)
alaska <- elide(alaska, shift=c(-2100000, -2500000))
proj4string(alaska) <- proj4string(us_counties_aea)

hawaii <- us_counties_aea[us_counties_aea$STATE=="15",]
hawaii <- elide(hawaii, rotate=-35)
hawaii <- elide(hawaii, shift=c(5400000, -1400000))
proj4string(hawaii) <- proj4string(us_counties_aea)

us_counties_aea <- us_counties_aea[!us_counties_aea$STATE %in% c("02", "15", "72"),]
us_counties_aea <- rbind(us_counties_aea, alaska, hawaii)

```

Finally, we tidy the spatial object into a data frame that ggplot can use, and clean up the `id` label by stripping out a prefix from the string. 

```{r, eval = FALSE, echo = TRUE, tidy = FALSE}
county_map <- tidy(us_counties_aea, region = "GEO_ID")
county_map$id <- stringr::str_replace(county_map$id,
                                      pattern = "0500000US", replacement = "")
```



```{r, eval = FALSE, echo = TRUE, tidy = FALSE}
county_map <- tidy(us_counties_aea, region = "GEO_ID")
county_map$id <- stringr::str_replace(county_map$id,
                                      pattern = "0500000US", replacement = "")
```


For more detail and code for the merge, see `github.com/kjhealy/us-county`.

## Plot themes and map themes

```{r, eval = FALSE}
devtools::install_github("hrbrmstr/hrbrthemes")
```


```{r, eval = FALSE, tidy = FALSE}

theme_map <- function(base_size=9, base_family="") {
    require(grid)
    theme_bw(base_size=base_size, base_family=base_family) %+replace%
        theme(axis.line=element_blank(),
              axis.text=element_blank(),
              axis.ticks=element_blank(),
              axis.title=element_blank(),
              panel.background=element_blank(),
              panel.border=element_blank(),
              panel.grid=element_blank(),
              panel.spacing=unit(0, "lines"),
              plot.background=element_blank(),
              legend.justification = c(0,0),
              legend.position = c(0,0)
              )
}

```
